---
title: "Visualize"
author: "Liz Ferguson"
date: "6/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## 3.1 Read Data

```{r}
# libraries
library(here)
library(readr)
library(DT)

# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

# read data
d_ac        <- read_csv(url_ac, col_names = F, skip = 2)
names(d_ac) <- names(read_csv(url_ac))

# show data
datatable(d_ac)
```


##3.2 Plot statically with ggplot2

###3.2.1 Simple line plot + geom_line()

```{r}
library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)


# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast


```


### 3.2.2 Trend line + geom_smooth()
```{r}
p_coast + 
  geom_smooth(method = "lm")

```


### 3.2.3 Distribution of values + geom_histogram()

```{r}
d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram()
```


### 3.2.4 Series line plot aes(color = region)

```{r}
library(stringr)
library(tidyr)

d_rgn <- d_ac %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # exclude column
  select(-total_fisheries_revenue_coastwide) %>% 
  # pivot longer (keeps "time" as is and pivots other columns)
  pivot_longer(-time) %>% 
  # mutate region by stripping other
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>% 
      str_to_upper()) %>% 
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)
  
# create plot object
p_rgn <- ggplot(
  d_rgn,
  # aesthetics
  aes(
    x     = time, 
    y     = value,
    group = region,
    color = region)) +
  # geometry
  geom_line()

# show plot
p_rgn

```


### 3.2.5 Update labels + labs()

```{r}
p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn
```


### 3.2.6 Multiple plots with facet_wrap()

```{r}
p_rgn +
  facet_wrap(vars(region))
```


### 3.2.7 Bar plot + geom_col()

# geom_bar doesn't allow you to indicate two variables, only uses x variables. 
# glue handles strings and nesting  and lubridate to use this year function
# x discrete (CA, OR, WA) and y is continuous
# use "fill" instead of "color" so it fills with color instead of just the outline 

```{r}
library(glue)
library(lubridate)

yr_max <- year(max(d_rgn$time))

d_rgn %>% 
  # filter by most recent time
  filter(year(time) == yr_max) %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_col() +
  # add labels
  labs(
    title  = glue("Fisheries Revenue for {yr_max}"),
    x      = "Region",
    y      = "Millions $ (year 2015)",
    fill   = "Region")
```


### 3.2.8 Variation of series with + geom_boxplot()

```{r}
d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_boxplot() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
```


### 3.2.9 Variation of series with + geom_violin()

```{r}
p_rgn_violin <- d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_violin() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
p_rgn_violin
```

### 3.2.10 Change Theme theme()

```{r}

p_rgn_violin +
  theme_classic()


# you can actually adjust to make your own theme
# first do theme_classic in console to see details
# Then adjust where you like (((here ))haven't done that to anything below, )only got rid of border here
theme_liz <- function (base_size = 11, base_family = "", base_line_size = base_size/22, 
    base_rect_size = base_size/22) 
{
    theme_bw(base_size = base_size, base_family = base_family, 
        base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace% 
        theme(#panel.border = element_blank(), 
              panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", 
                size = rel(1)), legend.key = element_blank(), 
            strip.background = element_rect(fill = "white", 
                colour = "black", size = rel(2)), complete = TRUE)
}



p_rgn_violin +
  # theme_classic()
  theme_liz()

```

## 3.3 Plot interactively with plotly or dygraphs
###3.3.1 Make ggplot interactive with plotly::ggplotly()

```{r}
# allows you to embed widgets in R markdown (only when html) - htmlwidgets for R 
# stores all of these into java libraries

plotly::ggplotly(p_rgn)

```


### 3.3.2 Create interactive time series with dygraphs::dygraph()
```{r}
library(dygraphs)

d_rgn_wide <- d_rgn %>% 
  mutate(
    Year = year(time)) %>% 
  select(Year, region, value) %>% 
  pivot_wider(
    names_from  = region,
    values_from = value)
datatable(d_rgn_wide)


d_rgn_wide %>% 
  dygraph() %>% 
  dyRangeSelector()
```

